---
title: "P8105 Homework 2"
author: "Connie Zhang"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

## Problem 1
* Importing trash wheel data and tidying/cleaning the data 
* Omitting non-data entries and rows that do not include dumpster specific data
```{r}
trash_wheel = 
  read_excel("./data/Trash-Wheel.xlsx", sheet = 1) %>%
  janitor::clean_names() %>%
  drop_na (dumpster) %>%
  select(-15, -16, -17) %>%
  mutate(sports_balls = as.integer(sports_balls))
trash_wheel 
```
* Read and clean precipitation data for 2017 
* Omitting rows without data and adding variable 2017
```{r}
precip_2017 =
  read_excel("./data/Trash-Wheel.xlsx", sheet = 6, range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate (year = 2017) %>%
  select(year,month,total)
precip_2017
```
* Read and clean precipitation data for 2018 
* Omitting rows without data and adding variable 2018
```{r}
precip_2018 =
  read_excel("./Data/Trash-Wheel.xlsx", sheet = 5, range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate (year = 2018) %>%
  select (year,month,total)
precip_2018
```
* Combining precipation datasets, converting month to character variable  
```{r}
combine_precip = 
  full_join(precip_2017,precip_2018) %>%
  mutate(month= as.numeric(month), month = month.name[month]) 
combine_precip
```
**Description of data:** 

The `trash_wheel` dataset contains information inlcuding amount of total litter collected and the type of litter collected, such as sports balls, plastic bags, and glass bottles .

The `combine_precip` dataset contains information on total precipitation for each month of 2017 and 2018 periods, measured in inches. 

There are `r dim(trash_wheel)[1]` observations in the `trash_wheel` dataset and `r dim(combine_precip)[1]` observations in the `combine_precip` dataset. 

The total amount of precipitation in 2018 was `r sum(pull(filter(combine_precip,year==2018),total))`. 

The median number of sports balls in a dumpster in 2017 is `r median(pull(filter(trash_wheel,year==2017),sports_balls))`. 

## Problem 2 
* Clean pols_month data and create a president variable 
```{r}
pols_month = 
  read_csv("./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into= c("year", "month", "day")) %>%
  mutate (month = as.numeric(month), month = month.abb[month], year= as.numeric(year), president = ifelse(prez_gop, "gop", "dem")) %>%
  select(-prez_gop, -prez_dem, -day)
pols_month
```
* Clean snp.csv data by arranging according to year and month
```{r}
snp_data = 
  read_csv("./data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, c("month", "day", "year")) %>%
  mutate(year = as.numeric(year), month = as.numeric(month), month = month.name[month]) %>%
  select(year, month, -day, close) %>%
  arrange (year, month)
snp_data

```
* Tidy unemployment.csv which will then be merged with snp.csv and pols-month.csv
```{r}
unemployment_data = 
  read_csv("./data/unemployment.csv") %>%
  pivot_longer((Jan:Dec), 
               names_to = "month", 
               values_to = "percent") %>%
  janitor::clean_names()
unemployment_data
```
* Merge snp.csv with pols-month.csv, and then with unemployment.csv
```{r}
polsnp_data =
 left_join(pols_month, snp_data, by = c("year", "month"))

merged_data = 
  left_join(polsnp_data, unemployment_data, by = c("year", "month"))
merged_data
```
**Description of data:** 

In the `pols_month` dataset, there were `r nrow(pols_month)` observations and `r ncol(pols_month)` variables. The years ranged from `r min(pull(pols_month, year))` to `r max(pull(pols_month, year))`. Key variables include specifying the elected president's presidential party affiliation, as well as the number of governors, senators and representatives across different time periods, with their party affiliation. 

In the `snp_data` dataset, there were `r nrow(snp_data)` observations and `r ncol(snp_data)` variables. The years ranged from `r min(pull(snp_data, year))` to `r max(pull(snp_data, year))`. Key variables include the closing values of the S&P stock index as they varied over different months of their respective years. 

In the `unemployment_data` dataset, there were `r nrow(unemployment_data)` observations and `r ncol(unemployment_data)` variables. The years ranged from `r min(pull(unemployment_data, year))` to `r max(pull(unemployment_data, year))`. Key variables include the unemployment rate across different months of their respective years. 

In the `merged_data` dataset, there were `r nrow(merged_data)` observations and `r ncol(merged_data)` variables. The years ranged from `r min(pull(merged_data, year))` to `r max(pull(merged_data, year))`. This includes all previously mentioned key variables into a combined dataset. 

## Problem 3
* Importing, loading, and tidying NYC Open data on popularity of baby names 
```{r}
baby_data =
  read_csv ("./data/popular_baby_names.csv") %>%
  janitor::clean_names() %>%
  distinct() %>%
  mutate(ethnicity =recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", "BLACK NON HISP" = "BLACK NON HISPANIC", "WHITE NON HISP" = "WHITE NON HISPANIC"), ethnicity = str_to_lower(ethnicity), gender = str_to_lower(gender), childs_first_name = str_to_lower (childs_first_name))
baby_data
```

* Creating table showing the name Olivia and its rank in popularity for female babies over time
```{r}
baby_data %>%
  filter (childs_first_name == "olivia") %>%
  select (-count) %>%
  pivot_wider(names_from = "year_of_birth", values_from = "rank") %>%
  knitr::kable()
baby_data
```

* Creating table showing the most popular male child name over time 
```{r}
baby_data %>%
  filter (gender == "male", rank == 1) %>%
  select (-count) %>%
  pivot_wider(names_from = "year_of_birth", values_from = "childs_first_name") %>%
  knitr::kable()
baby_data
```
* Creating scatter plot with popularity of name (x axis) againgst number of children with the name (y axis) for male, white non-hispanic children born in 2016. 

```{r}
scatterplot_child = baby_data %>% 
  filter(gender == "male", ethnicity == "white non hispanic", year_of_birth == "2016")
ggplot(scatterplot_child, aes(x = rank, y = count )) + geom_point()
```