---
title: "P8105 Homework 2"
author: "Connie Zhang"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

## Problem 1
```{r}
trash_wheel = 
  read_excel("./data/Trash-Wheel.xlsx", sheet = 1) %>%
  janitor::clean_names() %>%
  drop_na (dumpster) %>%
  select(-15, -16, -17) %>%
  mutate(sports_balls = as.integer(sports_balls))
  trash_wheel
```
* Read and clean precipitation data for 2017 
* Omitting rows without data and adding variable 2017
```{r}
precip_2017 =
  read_excel("./data/Trash-Wheel.xlsx", sheet = 6, range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate (year = 2017) %>%
  select(year,month,total)
precip_2017
```
* Read and clean precipitation data for 2018 
* Omitting rows without data and adding variable 2018
```{r}
precip_2018 =
  read_excel("./Data/Trash-Wheel.xlsx", sheet = 5, range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate (year = 2018) %>%
  select (year,month,total)
precip_2018
```
* Combining precipation datasets, converting month to character variable  
```{r}
combine_precip = 
  bind_rows(precip_2017,precip_2018) %>%
  mutate(month= as.numeric(month), month = month.name[month]) 
combine_precip
```
* **Description of data:** 
The Mr. Trash Wheel dataset contains information on dumpster number, date of trash collection, amount of total litter collected and the type of litter collected. 
The combined precipitation dataset contains information on total precipitation in inches for each month of 2017 and 2018 periods. 
The number of observations in both resulting datasets are `r NROW(pull(trash_wheel))` and `r NROW(pull(combine_precip))` respectively. 
The total amount of precipitation in 2018 was `r sum(pull(combine_precip,total))`.
The median number of sports balls in a dumpster in 2017 is `r median(pull(trash_wheel, sports_balls))`. 

## Problem 2